---
title: "Vantage Pro 2 Weather Station Data"
author: "Valeria Aspinall"
date: "2026-01-25"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# 2025 Data Upload, Reorganization & Identifying Gaps 
```{r}
library(knitr)
library(readr)
library(dplyr)
library(lubridate)
library(ggplot2)
library(plotly)
library(tidyverse)
library(data.table)
```

## 06-2025 Data Upload & Reorganization 
```{r}
#1. upload data using read.csv and the file path 
raw_jun25 <- read.csv("/Users/valeriaaspinall/Desktop/UGA Thesis Project /R Work - Thesis/Celeste Thesis/vantage_data/2025-06.csv", na = c("", "---"))

#2. merge the two metadata rows in one
hdr <- read_csv("/Users/valeriaaspinall/Desktop/UGA Thesis Project /R Work - Thesis/Celeste Thesis/vantage_data/2025-06.csv", na = c("", "---"), n_max = 2, col_names = FALSE)
col_names <- paste(hdr[1, ], hdr[2, ], sep = "_")
raw_jun25 <- read_csv("/Users/valeriaaspinall/Desktop/UGA Thesis Project /R Work - Thesis/Celeste Thesis/vantage_data/2025-06.csv", na = c("", "---"), skip = 2, col_names = col_names)

#3. mutate the date and time stamps into one column: data_jun25
data_jun25 <- raw_jun25 %>%
  mutate(DateTime = dmy_hms(paste(NA_Date, NA_Time))) %>%
  select(DateTime, everything(), -NA_Date, -NA_Time)

#4. ensuring timestamp is in the correct format 
data_jun25 <- data_jun25 %>% mutate(DateTime = as.POSIXct(DateTime))
```

## 06-2025 Identifying Gaps (Implicit & Explicit Missing Values) 
```{r}
#5. expanding data frame to identify implicit missing values 
##5a. starting by defining the start and end of the month of june 2025 
start_jun25 <- as.POSIXct("2025-06-01 00:15:00")
end_jun25   <- as.POSIXct("2025-07-01 00:00:00")

##5b. create the master sequence, 30 days, 15 min interval, should end up with 2880 observations 
jun25_full_timestamps <- seq(from = start_jun25, to = end_jun25, by = "15 min")

##5c. turn it into a master data frame: jun25_master_df
jun25_master_df <- data.frame(DateTime = jun25_full_timestamps)

##5d. make sure they are in the same time zone 
jun25_master_df <- jun25_master_df %>%
  mutate(DateTime = force_tz(DateTime, tzone = "America/Costa_Rica"))

data_jun25 <- data_jun25 %>%
  mutate(DateTime = force_tz(DateTime, tzone = "America/Costa_Rica"))

#6. merge with data_jun25: jun25_final to obtain the full range of june
jun25_final <- jun25_master_df %>%
  left_join(data_jun25, by = "DateTime")

##6a. check the final row number, for 30 days = 2880 
nrow(jun25_final)

#7. creating a heat map of missing values attempt 

##7a. deciding which variable to check, chose Out_Hum due to observation of it being lost first
variable_to_check <- "Out_Hum" 

jun25_gap_data <- jun25_final %>%
  mutate(is_missing = is.na(!!sym(variable_to_check))) %>%
  # Identify groups of consecutive TRUE (missing) values
  mutate(gap_id = rleid(is_missing)) %>%
  filter(is_missing == TRUE) %>%
  group_by(gap_id) %>%
  summarise(
    start_time = min(DateTime),
    end_time = max(DateTime),
    duration_hours = as.numeric(difftime(max(DateTime), min(DateTime), units = "hours")) + 0.25
  )

##7a. creating a ggplot of the time gaps 
jun25_gap_fig <- ggplot(jun25_gap_data) +
  # Draw a background line representing the whole month
  annotate("segment", x = start_jun25, xend = end_jun25, y = 1, yend = 1, 
           color = "grey80", size = 2) +
  # Overlay red segments for the gaps
  geom_segment(aes(x = start_time, xend = end_time, y = 1, yend = 1), 
               color = "red", size = 8) +
  scale_x_datetime(date_breaks = "2 days", date_labels = "%b %d") +
  theme_minimal() +
  theme(axis.text.y = element_blank(),
        axis.ticks.y = element_blank(),
        panel.grid.major.y = element_blank()) +
  labs(title = paste("Missing Data Intervals: 2025-06"),
       subtitle = "red bars indicate periods with no data",
       x = "June 2025", y = "Missing Gaps")

jun25_gap_fig
```


## 07-2025 Data Upload & Reorganization 
```{r}
#1. upload data using read.csv and the file path: raw_jul25
raw_jul25 <- read.csv("/Users/valeriaaspinall/Desktop/UGA Thesis Project /R Work - Thesis/Celeste Thesis/vantage_data/2025-07.csv", na = c("", "---"))

#2. merge the two metadata rows in one: raw_jul25
hdr <- read_csv("/Users/valeriaaspinall/Desktop/UGA Thesis Project /R Work - Thesis/Celeste Thesis/vantage_data/2025-07.csv", na = c("", "---"), n_max = 2, col_names = FALSE)
col_names <- paste(hdr[1, ], hdr[2, ], sep = "_")
raw_jul25 <- read_csv("/Users/valeriaaspinall/Desktop/UGA Thesis Project /R Work - Thesis/Celeste Thesis/vantage_data/2025-07.csv", na = c("", "---"), skip = 2, col_names = col_names)

#3. mutate the date and time stamps into one column: data_jul25
data_jul25 <- raw_jul25 %>%
  mutate(DateTime = dmy_hms(paste(NA_Date, NA_Time))) %>%
  select(DateTime, everything(), -NA_Date, -NA_Time)

#4. ensuring timestamp is in the correct format 
data_jul25 <- data_jul25 %>% mutate(DateTime = as.POSIXct(DateTime))
```

## 07-2025 Identifying Gaps (Implicit & Explicit Missing Values) 
```{r}
#5. expanding data frame to identify implicit missing values 
##5a. starting by defining the start and end of the month of july 2025 
start_jul25 <- as.POSIXct("2025-07-01 00:15:00")
end_jul25   <- as.POSIXct("2025-08-01 00:00:00")

##5b. create the master sequence, 31 days, 15 min interval, should end up with 2976 observations 
jul25_full_timestamps <- seq(from = start_jul25, to = end_jul25, by = "15 min")

##5c. turn it into a master data frame: jul25_master_df
jul25_master_df <- data.frame(DateTime = jul25_full_timestamps)

##5d. make sure they are in the same time zone 
jul25_master_df <- jul25_master_df %>%
  mutate(DateTime = force_tz(DateTime, tzone = "America/Costa_Rica"))

data_jul25 <- data_jul25 %>%
  mutate(DateTime = force_tz(DateTime, tzone = "America/Costa_Rica"))

#6. merge with data_jul25: jul25_final
jul25_final <- jul25_master_df %>%
  left_join(data_jul25, by = "DateTime")

##6a. check the final row number, for 30 days = 2880 
nrow(jul25_final)
head(jul25_final)

#7. creating a heat map of missing values attempt 

##7a. deciding which variable to check, chose Out_Hum due to observation of it being lost first
variable_to_check <- "Out_Hum" 

jul25_gap_data <- jul25_final %>%
  mutate(is_missing = is.na(!!sym(variable_to_check))) %>%
  # Identify groups of consecutive TRUE (missing) values
  mutate(gap_id = rleid(is_missing)) %>%
  filter(is_missing == TRUE) %>%
  group_by(gap_id) %>%
  summarise(
    start_time = min(DateTime),
    end_time = max(DateTime),
    duration_hours = as.numeric(difftime(max(DateTime), min(DateTime), units = "hours")) + 0.25
  )

##7a. creating a ggplot of the time gaps 
jul25_gap_fig <- ggplot(jul25_gap_data) +
  # Draw a background line representing the whole month
  annotate("segment", x = start_jul25, xend = end_jul25, y = 1, yend = 1, 
           color = "grey80", size = 2) +
  # Overlay red segments for the gaps
  geom_segment(aes(x = start_time, xend = end_time, y = 1, yend = 1), 
               color = "red", size = 8) +
  scale_x_datetime(date_breaks = "2 days", date_labels = "%b %d") +
  theme_minimal() +
  theme(axis.text.y = element_blank(),
        axis.ticks.y = element_blank(),
        panel.grid.major.y = element_blank()) +
  labs(title = paste("Missing Data Intervals: 2025-07"),
       subtitle = "red bars indicate periods with no data",
       x = "July 2025", y = "Missing Gaps")

jul25_gap_fig
```


## 08-2025 Data Upload & Reorganization 
```{r}
#1. upload data using read.csv and the file path: raw_ago25
raw_ago25 <- read.csv("/Users/valeriaaspinall/Desktop/UGA Thesis Project /R Work - Thesis/Celeste Thesis/vantage_data/2025-08.csv", na = c("", "---"))

#2. merge the two metadata rows in one: raw_ago25
hdr <- read_csv("/Users/valeriaaspinall/Desktop/UGA Thesis Project /R Work - Thesis/Celeste Thesis/vantage_data/2025-08.csv", na = c("", "---"), n_max = 2, col_names = FALSE)
col_names <- paste(hdr[1, ], hdr[2, ], sep = "_")
raw_ago25 <- read_csv("/Users/valeriaaspinall/Desktop/UGA Thesis Project /R Work - Thesis/Celeste Thesis/vantage_data/2025-08.csv", na = c("", "---"), skip = 2, col_names = col_names)

#3. mutate the date and time stamps into one column: data_ago25
data_ago25 <- raw_ago25 %>%
  mutate(DateTime = dmy_hms(paste(NA_Date, NA_Time))) %>%
  select(DateTime, everything(), -NA_Date, -NA_Time)

#4. ensuring timestamp is in the correct format 
data_ago25 <- data_ago25 %>% mutate(DateTime = as.POSIXct(DateTime))

head(data_ago25)
tail(data_ago25)
nrow(data_ago25)
```

## 08-2025 Identifying Gaps (Implicit & Explicit Missing Values) 
```{r}
#5. expanding data frame to identify implicit missing values 
##5a. starting by defining the start and end of the month of august 2025 
start_ago25 <- as.POSIXct("2025-08-01 00:15:00")
end_ago25   <- as.POSIXct("2025-09-01 00:00:00")

##5b. create the master sequence, 31 days, 15 min interval, should end up with 2976 observations 
ago25_full_timestamps <- seq(from = start_ago25, to = end_ago25, by = "15 min")

##5c. turn it into a master data frame: ago25_master_df
ago25_master_df <- data.frame(DateTime = ago25_full_timestamps)

##5d. make sure they are in the same time zone 
ago25_master_df <- ago25_master_df %>%
  mutate(DateTime = force_tz(DateTime, tzone = "America/Costa_Rica"))

data_ago25 <- data_ago25 %>%
  mutate(DateTime = force_tz(DateTime, tzone = "America/Costa_Rica"))

#6. merge with data_ago25: ago25_final
ago25_final <- ago25_master_df %>%
  left_join(data_ago25, by = "DateTime")

##6a. check the final row number 
nrow(ago25_final)
head(ago25_final)

#7. creating a heat map of missing values attempt 

##7a. deciding which variable to check, chose Out_Hum due to observation of it being lost first
variable_to_check <- "Out_Hum" 

ago25_gap_data <- ago25_final %>%
  mutate(is_missing = is.na(!!sym(variable_to_check))) %>%
  # Identify groups of consecutive TRUE (missing) values
  mutate(gap_id = rleid(is_missing)) %>%
  filter(is_missing == TRUE) %>%
  group_by(gap_id) %>%
  summarise(
    start_time = min(DateTime),
    end_time = max(DateTime),
    duration_hours = as.numeric(difftime(max(DateTime), min(DateTime), units = "hours")) + 0.25
  )

##7a. creating a ggplot of the time gaps 
ago25_gap_fig <- ggplot(ago25_gap_data) +
  # Draw a background line representing the whole month
  annotate("segment", x = start_ago25, xend = end_ago25, y = 1, yend = 1, 
           color = "grey80", size = 2) +
  # Overlay red segments for the gaps
  geom_segment(aes(x = start_time, xend = end_time, y = 1, yend = 1), 
               color = "red", size = 8) +
  scale_x_datetime(date_breaks = "2 days", date_labels = "%b %d") +
  theme_minimal() +
  theme(axis.text.y = element_blank(),
        axis.ticks.y = element_blank(),
        panel.grid.major.y = element_blank()) +
  labs(title = paste("Missing Data Intervals: 2025-08"),
       subtitle = "red bars indicate periods with no data",
       x = "August 2025", y = "Missing Gaps")

ago25_gap_fig
```


## 09-2025 Data Upload & Reorganization 
```{r}
#1. upload data using read.csv and the file path: raw_set25
raw_sep25 <- read.csv("/Users/valeriaaspinall/Desktop/UGA Thesis Project /R Work - Thesis/Celeste Thesis/vantage_data/2025-09.csv", na = c("", "---"))

#2. merge the two metadata rows in one: raw_set25
hdr <- read_csv("/Users/valeriaaspinall/Desktop/UGA Thesis Project /R Work - Thesis/Celeste Thesis/vantage_data/2025-09.csv", na = c("", "---"), n_max = 2, col_names = FALSE)
col_names <- paste(hdr[1, ], hdr[2, ], sep = "_")
raw_sep25 <- read_csv("/Users/valeriaaspinall/Desktop/UGA Thesis Project /R Work - Thesis/Celeste Thesis/vantage_data/2025-09.csv", na = c("", "---"),skip = 2, col_names = col_names)

#3. mutate the date and time stamps into one column: data_set25
data_sep25 <- raw_sep25 %>%
  mutate(DateTime = dmy_hms(paste(NA_Date, NA_Time))) %>%
  select(DateTime, everything(), -NA_Date, -NA_Time)

#4. ensuring timestamp is in the correct format 
data_sep25 <- data_sep25 %>% mutate(DateTime = as.POSIXct(DateTime))
```

## 09-2025 Identifying Gaps (Implicit & Explicit Missing Values) 
```{r}
#5. expanding data frame to identify implicit missing values 
##5a. starting by defining the start and end of the month of september 2025 
start_sep25 <- as.POSIXct("2025-09-01 00:15:00")
end_sep25   <- as.POSIXct("2025-10-01 00:00:00")

##5b. create the master sequence, 30 days, 15 min interval, should end up with 2880 observations 
sep25_full_timestamps <- seq(from = start_sep25, to = end_sep25, by = "15 min")

##5c. turn it into a master data frame: sep25_master_df
sep25_master_df <- data.frame(DateTime = sep25_full_timestamps)

##5d. make sure they are in the same time zone 
sep25_master_df <- sep25_master_df %>%
  mutate(DateTime = force_tz(DateTime, tzone = "America/Costa_Rica"))

data_sep25 <- data_sep25 %>%
  mutate(DateTime = force_tz(DateTime, tzone = "America/Costa_Rica"))

#6. merge with data_sep25: sep25_final
sep25_final <- sep25_master_df %>%
  left_join(data_sep25, by = "DateTime")

##6a. check the final row number, for 30 days = 2880 
nrow(sep25_final)

#7. creating a heat map of missing values attempt 

##7a. deciding which variable to check, chose Out_Hum due to observation of it being lost first
variable_to_check <- "Out_Hum" 

sep25_gap_data <- sep25_final %>%
  mutate(is_missing = is.na(!!sym(variable_to_check))) %>%
  # Identify groups of consecutive TRUE (missing) values
  mutate(gap_id = rleid(is_missing)) %>%
  filter(is_missing == TRUE) %>%
  group_by(gap_id) %>%
  summarise(
    start_time = min(DateTime),
    end_time = max(DateTime),
    duration_hours = as.numeric(difftime(max(DateTime), min(DateTime), units = "hours")) + 0.25
  )

##7a. creating a ggplot of the time gaps 
sep25_gap_fig <- ggplot(sep25_gap_data) +
  # Draw a background line representing the whole month
  annotate("segment", x = start_sep25, xend = end_sep25, y = 1, yend = 1, 
           color = "grey80", size = 2) +
  # Overlay red segments for the gaps
  geom_segment(aes(x = start_time, xend = end_time, y = 1, yend = 1), 
               color = "red", size = 8) +
  scale_x_datetime(date_breaks = "2 days", date_labels = "%b %d") +
  theme_minimal() +
  theme(axis.text.y = element_blank(),
        axis.ticks.y = element_blank(),
        panel.grid.major.y = element_blank()) +
  labs(title = paste("Missing Data Intervals: 2025-09"),
       subtitle = "red bars indicate periods with no data",
       x = "September 2025", y = "Missing Gaps")

sep25_gap_fig
```


## 10-2025 Data Upload & Reorganization 
```{r}
#1. upload data using read.csv and the file path: raw_oct25
raw_oct25 <- read.csv("/Users/valeriaaspinall/Desktop/UGA Thesis Project /R Work - Thesis/Celeste Thesis/vantage_data/2025-10.csv", na = c("", "---"))

#2. merge the two metadata rows in one: raw_oct25
hdr <- read_csv("/Users/valeriaaspinall/Desktop/UGA Thesis Project /R Work - Thesis/Celeste Thesis/vantage_data/2025-10.csv", na = c("", "---"), n_max = 2, col_names = FALSE)
col_names <- paste(hdr[1, ], hdr[2, ], sep = "_")
raw_oct25 <- read_csv("/Users/valeriaaspinall/Desktop/UGA Thesis Project /R Work - Thesis/Celeste Thesis/vantage_data/2025-10.csv", na = c("", "---"), skip = 2, col_names = col_names)

#3. mutate the date and time stamps into one column: data_oct25
data_oct25 <- raw_oct25 %>%
  mutate(DateTime = dmy_hms(paste(NA_Date, NA_Time))) %>%
  select(DateTime, everything(), -NA_Date, -NA_Time)

#4. ensuring timestamp is in the correct format 
data_oct25 <- data_oct25 %>% mutate(DateTime = as.POSIXct(DateTime))
```

## 10-2025 Identifying Gaps (Implicit & Explicit Missing Values)
```{r}
#5. expanding data frame to identify implicit missing values 
##5a. starting by defining the start and end of the month of oct 2025 
start_oct25 <- as.POSIXct("2025-10-01 00:15:00")
end_oct25   <- as.POSIXct("2025-11-01 00:00:00")

##5b. create the master sequence, 31 days, 15 min interval, should end up with 2976 observations 
oct25_full_timestamps <- seq(from = start_oct25, to = end_oct25, by = "15 min")

##5c. turn it into a master data frame: oct25_master_df
oct25_master_df <- data.frame(DateTime = oct25_full_timestamps)

##5d. make sure they are in the same time zone 
oct25_master_df <- oct25_master_df %>%
  mutate(DateTime = force_tz(DateTime, tzone = "America/Costa_Rica"))

data_oct25 <- data_oct25 %>%
  mutate(DateTime = force_tz(DateTime, tzone = "America/Costa_Rica"))

#6. merge with data_oct25: oct25_final
oct25_final <- oct25_master_df %>%
  left_join(data_oct25, by = "DateTime")

##6a. check the final row number, for 31 days = 2976 
nrow(oct25_final)

#7. creating a heat map of missing values attempt 

##7a. deciding which variable to check, chose Out_Hum due to observation of it being lost first
variable_to_check <- "Out_Hum" 

oct25_gap_data <- oct25_final %>%
  mutate(is_missing = is.na(!!sym(variable_to_check))) %>%
  # Identify groups of consecutive TRUE (missing) values
  mutate(gap_id = rleid(is_missing)) %>%
  filter(is_missing == TRUE) %>%
  group_by(gap_id) %>%
  summarise(
    start_time = min(DateTime),
    end_time = max(DateTime),
    duration_hours = as.numeric(difftime(max(DateTime), min(DateTime), units = "hours")) + 0.25
  )

##7a. creating a ggplot of the time gaps 
oct25_gap_fig <- ggplot(oct25_gap_data) +
  # Draw a background line representing the whole month
  annotate("segment", x = start_oct25, xend = end_oct25, y = 1, yend = 1, 
           color = "grey80", size = 2) +
  # Overlay red segments for the gaps
  geom_segment(aes(x = start_time, xend = end_time, y = 1, yend = 1), 
               color = "red", size = 8) +
  scale_x_datetime(date_breaks = "2 days", date_labels = "%b %d") +
  theme_minimal() +
  theme(axis.text.y = element_blank(),
        axis.ticks.y = element_blank(),
        panel.grid.major.y = element_blank()) +
  labs(title = paste("Missing Data Intervals: 2025-10"),
       subtitle = "red bars indicate periods with no data",
       x = "October 2025", y = "Missing Gaps")

oct25_gap_fig
```


## 11-2025 Data Upload & Reorganization 
```{r}
#1. upload data using read.csv and the file path: raw_nov25
raw_nov25 <- read.csv("/Users/valeriaaspinall/Desktop/UGA Thesis Project /R Work - Thesis/Celeste Thesis/vantage_data/2025-11.csv", na = c("", "---"))

#2. merge the two metadata rows in one: raw_nov25
hdr <- read_csv("/Users/valeriaaspinall/Desktop/UGA Thesis Project /R Work - Thesis/Celeste Thesis/vantage_data/2025-11.csv", na = c("", "---"), n_max = 2, col_names = FALSE)
col_names <- paste(hdr[1, ], hdr[2, ], sep = "_")
raw_nov25 <- read_csv("/Users/valeriaaspinall/Desktop/UGA Thesis Project /R Work - Thesis/Celeste Thesis/vantage_data/2025-11.csv", na = c("", "---"), skip = 2, col_names = col_names)

#3. mutate the date and time stamps into one column: data_nov25
data_nov25 <- raw_nov25 %>%
  mutate(DateTime = dmy_hms(paste(NA_Date, NA_Time))) %>%
  select(DateTime, everything(), -NA_Date, -NA_Time)

#4. ensuring timestamp is in the correct format 
data_nov25 <- data_nov25 %>% mutate(DateTime = as.POSIXct(DateTime))
```

## 11-2025 Identifying Gaps (Implicit & Explicit Missing Values)
```{r}
#5. expanding data frame to identify implicit missing values 
##5a. starting by defining the start and end of the month of nov 2025 
start_nov25 <- as.POSIXct("2025-11-01 00:15:00", tz = "America/Costa_Rica")
end_nov25   <- as.POSIXct("2025-12-01 00:00:00", tz = "America/Costa_Rica")

##5b. create the master sequence, 30 days, 15 min interval, should end up with 2880 observations 
nov25_full_timestamps <- seq(from = start_nov25, to = end_nov25, by = "15 min")

##5c. turn it into a master data frame: nov25_master_df
nov25_master_df <- data.frame(DateTime = nov25_full_timestamps)
nrow(nov25_master_df)

##5d. make sure they are in the same time zone 
nov25_master_df <- nov25_master_df %>%
  mutate(DateTime = force_tz(DateTime, tzone = "America/Costa_Rica"))

data_nov25 <- data_nov25 %>%
  mutate(DateTime = force_tz(DateTime, tzone = "America/Costa_Rica"))

nrow(data_nov25)
head(data_nov25)

#6. merge with data_nov25: nov25_final
nov25_final <- nov25_master_df %>%
  left_join(data_nov25, by = "DateTime")

##6a. check the final row number, for 31 days = 2976 
nrow(nov25_final)

#7. creating a heat map of missing values attempt 

##7a. deciding which variable to check, chose Out_Hum due to observation of it being lost first
variable_to_check <- "Out_Hum" 

nov25_gap_data <- nov25_final %>%
  mutate(is_missing = is.na(!!sym(variable_to_check))) %>%
  # Identify groups of consecutive TRUE (missing) values
  mutate(gap_id = rleid(is_missing)) %>%
  filter(is_missing == TRUE) %>%
  group_by(gap_id) %>%
  summarise(
    start_time = min(DateTime),
    end_time = max(DateTime),
    duration_hours = as.numeric(difftime(max(DateTime), min(DateTime), units = "hours")) + 0.25
  )

##7a. creating a ggplot of the time gaps 
nov25_gap_fig <- ggplot(nov25_gap_data) +
  # Draw a background line representing the whole month
  annotate("segment", x = start_nov25, xend = end_nov25, y = 1, yend = 1, 
           color = "grey80", size = 2) +
  # Overlay red segments for the gaps
  geom_segment(aes(x = start_time, xend = end_time, y = 1, yend = 1), 
               color = "red", size = 8) +
  scale_x_datetime(date_breaks = "2 days", date_labels = "%b %d") +
  theme_minimal() +
  theme(axis.text.y = element_blank(),
        axis.ticks.y = element_blank(),
        panel.grid.major.y = element_blank()) +
  labs(title = paste("Missing Data Intervals: 2025-11"),
       subtitle = "red bars indicate periods with no data",
       x = "November 2025", y = "Missing Gaps")

nov25_gap_fig
```


## 12-2025 Data Upload & Reorganization 
```{r}
#1. upload data using read.csv and the file path: raw_dec25
raw_dec25 <- read.csv("/Users/valeriaaspinall/Desktop/UGA Thesis Project /R Work - Thesis/Celeste Thesis/vantage_data/2025-12.csv", na = c("", "---"))

#2. merge the two metadata rows in one: raw_dec25
hdr <- read_csv("/Users/valeriaaspinall/Desktop/UGA Thesis Project /R Work - Thesis/Celeste Thesis/vantage_data/2025-12.csv", na = c("", "---"), n_max = 2, col_names = FALSE)
col_names <- paste(hdr[1, ], hdr[2, ], sep = "_")
raw_dec25 <- read_csv("/Users/valeriaaspinall/Desktop/UGA Thesis Project /R Work - Thesis/Celeste Thesis/vantage_data/2025-12.csv", na = c("", "---"),skip = 2, col_names = col_names)

#3. mutate the date and time stamps into one column: data_nov25
data_dec25 <- raw_dec25 %>%
  mutate(DateTime = dmy_hms(paste(NA_Date, NA_Time))) %>%
  select(DateTime, everything(), -NA_Date, -NA_Time)

#4. ensuring timestamp is in the correct format 
data_dec25 <- data_dec25 %>% mutate(DateTime = as.POSIXct(DateTime))
```

## 12-2025 Identifying Gaps (Implicit & Explicit Missing Values)
```{r}
#5. expanding data frame to identify implicit missing values 
##5a. starting by defining the start and end of the month of dec 2025 
start_dec25 <- as.POSIXct("2025-12-01 00:15:00")
end_dec25   <- as.POSIXct("2026-01-01 00:00:00")

##5b. create the master sequence, 31 days, 15 min interval, should end up with 2976 observations 
dec25_full_timestamps <- seq(from = start_dec25, to = end_dec25, by = "15 min")

##5c. turn it into a master data frame: dec25_master_df
dec25_master_df <- data.frame(DateTime = dec25_full_timestamps)

##5d. make sure they are in the same time zone 
dec25_master_df <- dec25_master_df %>%
  mutate(DateTime = force_tz(DateTime, tzone = "America/Costa_Rica"))

data_dec25 <- data_dec25 %>%
  mutate(DateTime = force_tz(DateTime, tzone = "America/Costa_Rica"))

head(data_dec25)

#6. merge with data_nov25: nov25_final
dec25_final <- dec25_master_df %>%
  left_join(data_dec25, by = "DateTime")

##6a. check the final row number, for 31 days = 2976 
nrow(dec25_final)

#7. creating a heat map of missing values attempt 

##7a. deciding which variable to check, chose Out_Hum due to observation of it being lost first
variable_to_check <- "Out_Hum" 

dec25_gap_data <- dec25_final %>%
  mutate(is_missing = is.na(!!sym(variable_to_check))) %>%
  # Identify groups of consecutive TRUE (missing) values
  mutate(gap_id = rleid(is_missing)) %>%
  filter(is_missing == TRUE) %>%
  group_by(gap_id) %>%
  summarise(
    start_time = min(DateTime),
    end_time = max(DateTime),
    duration_hours = as.numeric(difftime(max(DateTime), min(DateTime), units = "hours")) + 0.25
  )

##7a. creating a ggplot of the time gaps 
dec25_gap_fig <- ggplot(dec25_gap_data) +
  # Draw a background line representing the whole month
  annotate("segment", x = start_dec25, xend = end_dec25, y = 1, yend = 1, 
           color = "grey80", size = 2) +
  # Overlay red segments for the gaps
  geom_segment(aes(x = start_time, xend = end_time, y = 1, yend = 1), 
               color = "red", size = 8) +
  scale_x_datetime(date_breaks = "2 days", date_labels = "%b %d") +
  theme_minimal() +
  theme(axis.text.y = element_blank(),
        axis.ticks.y = element_blank(),
        panel.grid.major.y = element_blank()) +
  labs(title = paste("Missing Data Intervals: 2025-12"),
       subtitle = "red bars indicate periods with no data",
       x = "Decemeber 2025", y = "Missing Gaps")

dec25_gap_fig
```





# Merging 2025 Data together 
```{r}
full_2025_data <- bind_rows(jun25_final,jul25_final,ago25_final,sep25_final,oct25_final,nov25_final,dec25_final)

full_2025_data <- full_2025_data %>%
  arrange(DateTime)

nrow(full_2025_data)
```

# 06-2025 - Plotly Trials w/ Rain
```{r}
#1. trying plotly on the range of june-2025 to decemeber-2025 with precipitation 

fig_2025_rain <- plot_ly(width = 900, height = 500) %>%
  add_trace(data = full_2025_data, x = ~`DateTime`, y = ~ `NA_Rain`,
            name = '2025 Precipitation', type = 'scatter', mode = 'lines',
            line = list(width = 0.8))
fig_2025_rain

#2. now trying plotly but with daily sum of precipitation 

##2a. first, obtain the daily precipitation values with the following code
daily_precip <- full_2025_data %>%
  # create a 'date' column that ignores the 15-min time
  mutate(Date = as.Date(DateTime)) %>%
  # group by that date
  group_by(Date) %>%
  # calculate the sum precipitation for each day
  # Use na.rm = TRUE to ignore missing values in the average
  summarise(Daily_Precip = sum(NA_Rain, na.rm = TRUE))

##2b. now, attempt using the plotly again 

fig25_daily_precip <- plot_ly(width = 900, height = 500) %>%
  add_trace(data = daily_precip, x = ~`Date`, y = ~ `Daily_Precip`,
            name = '2025 Precipitation', type = 'scatter', mode = 'lines', fill = 'tozeroy',
            line = list(width = 0.8))

fig25_daily_precip
```





